\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-12 icra-2021.tex
cleaned and added an appendix

%% pseudocode package
%\usepackage{algorithmic}
\usepackage{algpseudocode}

@@

@@

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{&\langle\mathbf{c}_{k},\mathbf{s}_{k}\rangle\mid\,\exists n\in\mathbb{Z}_{>0},\\&\langle\varphi_{i-1}(\mathbf{p}_{k-n},\mathbf{c}_{k-n}),\Psi(\mathbf{s}_{k-n})\rangle\in\mathcal{M}_{i-1}\\
    &\Longrightarrow\langle\varphi_i(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k})\rangle\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}

@@

\begin{figure*}[h]
  \centering
  \hspace{-5ex}\subfigure[.]
  {
    \input{\figpath/energy/est_vs_joules.tikz}
    \label{fig:energy-max}
  }
  \subfigure[.]
  {
    \input{\figpath/energy/max_qos_tees3.tikz}
    \label{fig:energy-max}
  }
  \caption{.}
\end{figure*}

@@

In this section

\begin{figure}[h]
  \centering
  \input{\figpath/trajectory/trajectory3.tikz}
  \label{fig:trajectory}
  \caption{.}
\end{figure}

@@

\bibliographystyle{IEEEtran}
\bibliography{\jobname} 
\vspace{0.1ex}

\newpage

>>>>

%% pseudocode package
\usepackage{algpseudocode}

@@

% just lines in pseudocode...
\usepackage{etoolbox}

\makeatletter
% start with some helper code
% This is the vertical rule that is inserted
\newcommand*{\algrule}[1][\algorithmicindent]{%
  \makebox[#1][l]{%
    \hspace*{.2em}% <------------- This is where the rule starts from
    \vrule height .75\baselineskip depth .25\baselineskip
  }
}

\newcount\ALG@printindent@tempcnta
\def\ALG@printindent{%
    \ifnum \theALG@nested>0% is there anything to print
    \ifx\ALG@text\ALG@x@notext% is this an end group without any text?
    % do nothing
    \else
    \unskip
    % draw a rule for each indent level
    \ALG@printindent@tempcnta=1
    \loop
    \algrule[\csname ALG@ind@\the\ALG@printindent@tempcnta\endcsname]%
    \advance \ALG@printindent@tempcnta 1
    \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax
    \repeat
    \fi
    \fi
}
% the following line injects our new indent handling code in place of the default spacing
\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}{\ALG@printindent}{}{\errmessage{failed to patch}}
\patchcmd{\ALG@doentity}{\item[]\nointerlineskip}{}{}{} % no spurious vertical space
% end vertical rule patch for algorithmicx
\makeatother

% for blank page

\usepackage{afterpage}

@@

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}\label{eq:mission-valid}
    \mathbf{u}^a_{k}=\{&\langle\mathbf{c}_{k},\mathbf{s}_{k}\rangle\mid\,\exists n\in\mathbb{Z}_{>0},\\&\langle\varphi_{i-1}(\mathbf{p}_{k-n},\mathbf{c}_{k-n}),\Psi(\mathbf{s}_{k-n})\rangle\in\mathcal{M}_{i-1}\\
    &\Longrightarrow\langle\varphi_i(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k})\rangle\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}

@@

@@

Some preliminary results are discussed. Further analysis and re-implementation of the functionality of the algorithm is needed.

\begin{figure}[h]
  \centering
  \input{\figpath/trajectory/trajectory3.tikz}
  \caption{The true trajectory. The trajectory is retrieved from the flight log.}
  \label{fig:trajectory}
\end{figure}

\begin{figure}[h]
  \centering
  \input{\figpath/energy/max_qos_tees3.tikz}
  \caption{Derived energy consumption from the sensors (throttle). Red line is the estimated value. We use KF for state estimation when the difference between the sensor and model value is $>\varepsilon$. Otherwise we evolve the state without noise.}
  \label{fig:energy-max}
\end{figure}

@@

\leavevmode\thispagestyle{empty}\newpage % just for now

\bibliographystyle{IEEEtran}
\bibliography{\jobname}

\leavevmode\thispagestyle{empty}\newpage

\appendices
\section{Mission Plan Format}

In this appendix we discuss how to specify an initial mission plan with an example of a survey mission. The reader can later execute the initial mission plan using the simulator in Appendix~\ref{app:simulator}. The initial mission plan is user-defined. It specifies the trajectory and tasks of a mission along the QoS and TEEs sets.

Recall that the mission is composed of a set of stages (Subsection~\ref{sec:prob}). The plan has a variable number of entries per each stage. Suppose we want to plan the mission in Figures~\ref{fig:overview},~\ref{fig:trajectory}. The first line
\begin{algorithmic}[1]
  \State\textbf{\texttt{0.0003;90;(x+45)\^{}2+(y-146)\^{}2-4900+c1}}\label{code:firstcircle}
\end{algorithmic}
corresponds to TEE $\varphi_0(\mathbf{p}_k,c_{0,1}):=(x+45)^2+(y-146)^2-4900+c_{0,1}$ in the initial stage $\mathcal{M}_0$. The TEE is a circle. The gain is $k_e=3\cdot 10^{-4}$, and the rotation is of ninety degrees--we use the rotation matrix $E$ from Equation~(\ref{eq:pd}). Note that line~\ref{code:firstcircle} further specifies the adjustment {\tt\textbf{c1}} of the radius of the circle. It corresponds to $c_{0,1}$ in Definition~\ref{def:mission} (zero numerates the stage, one the adjustment).

The following line
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:firstcircle}
  \State\textbf{\texttt{-sqrt(4900+c1)-45,146}}\label{code:trigger}
\end{algorithmic}
specifies the triggering point. {\tt\textbf{sqrt}} is the square root. The UAV reaches the triggering point within a given $\varepsilon$ (the tolerance), and the algorithm switches to the next stage $\mathcal{M}_1$. The algorithm ensures that the overall mission is still valid--Equation~(\ref{eq:mission-valid}). It ensure that the TEEs and QoS constraints sets are satisfied with the generated control input. Note that the triggering point is in function of the adjustment {\tt\textbf{c1}}. This is necessary, an adjustment in one stage affects the overall flight.

Let us assume the system has two tasks. One task is the CNN object detection algorithm. Its computation $s_{0,1}$ is the fps rate. Another task is the variable key-size encryption algorithm. Its computation $s_{0,2}$ is the key-size. Then in mission plan
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:trigger}
  \State\textbf{\texttt{[-3*10\^{}3,0]}}\label{code:teeset}
  \State\textbf{\texttt{[2,10]}}\label{code:qos1set}
  \State\textbf{\texttt{[32,448]}}\label{code:qos2set}
\end{algorithmic}
the line~\ref{code:teeset} corresponds to TEE $\varphi_0$'s set $\mathbb{C}_0=[-3\cdot10^3,0]$. The algorithm selects $c_{0,1}$ in $\mathbb{C}_0$. The line~\ref{code:qos1set} corresponds to the QoS set $\mathbb{S}_{0,1}$ for the computation $s_{0,1}$. The line~\ref{code:qos1set} corresponds to the QoS set $\mathbb{S}_{0,2}$ for the computation $s_{0,2}$. The lines~\ref{code:teeset}--\ref{code:qos2set} must follow the ascending order of adjustments and computations. The lines~\ref{code:firstcircle}--\ref{code:qos2set} all describe the stage $\mathcal{M}_0$.

Once the UAV reaches the triggering point (line~\ref{code:trigger}) and there are no further lines, the algorithm terminates. If there are further lines, the stage switches to $\mathcal{M}_1$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:qos2set}
  \State\textbf{\texttt{0.05;270;x+sqrt(4900+c1)+45}}\label{code:tee2}
  \State\textbf{\texttt{-sqrt(4900+c1)-45,11}}\label{code:trigger2}
  \State\textbf{\texttt{[-3*10\^{}3,0]}}\label{code:teeset2}
  \State\textbf{\texttt{[2,10]}}
  \State\textbf{\texttt{[32,448]}}\label{code:qosset12}
\end{algorithmic}
line~\ref{code:tee2} corresponds to TEE $\varphi_1(\mathbf{p}_k,c_{0,1}):=x+\sqrt{4900+c_{0,1}}+45$, a linear equation that intersects the triggering point (line~\ref{code:trigger}). The gain is $k_e=5\cdot 10^{-2}$, and the rotation is opposite to line~\ref{code:firstcircle}. The algorithm computes $-E$ from the rotation matrix in Equation~\ref{eq:pd}.

The gain is different from the one used in $\varphi_0$ as different equations require different convergence rate. When we follow a circle it is desirable to turn the rotation direction in advance. When we follow a straight line, it is more desirable to head to the line and turn the rotation direction while closer to the line. We refer to the simulator in Appendix~\ref{app:simulator} for more details on how to tune the gain $k_e$. The reader can simulate different rates and see how they affect the flight. The rotation is likewise different. In fact, the rotation matrix $E$ points upwards in the space and heads to the counter-clockwise direction. The rotation matrix $-E$ points downwards and head to the clockwise direction.

The triggering point of the stage $\mathcal{M}_1$ at line~\ref{code:trigger2} also depends on the adjustment {\tt\textbf{c1}}. The lines~\ref{code:teeset2}--\ref{code:qosset12} specifies the constraints sets. They are expressed the same way as lines~\ref{code:teeset}--\ref{code:qos2set}. They have to be specified explicitly for each stage, although they don't change. The current implementation of the algorithm has no other means to distinguish different constraints sets.

Again, if there are further lines and UAV reaches the triggering point (line~\ref{code:trigger2}), the stage switches to $\mathcal{M}_2$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:qosset12}
  \State\textbf{\texttt{0.0003;90;(x+sqrt(4900+c1)-30)\^{}2}}
  \textbf{\texttt{\hspace*{14ex}+(y-11)\^{}2-5625}}\label{code:tee3}
  \State\textbf{\texttt{-sqrt(4900+c1)+105,11}}
  \State\textbf{\texttt{[-3*10\^{}3,0]}}
  \State\textbf{\texttt{[2,10]}}
  \State\textbf{\texttt{[32,448]}}\label{code:qosset32}
\end{algorithmic}
line~\ref{code:tee3} corresponds to TEE $\varphi_2(\mathbf{p}_k,c_{0,1}):=(x+\sqrt{4900+c_{0,1}}-30)^2+(y-11)^2-5625$, a circle of fixed size that changes the coordinates in function of the adjustment {\tt\textbf{c1}}. The rotation is counter-clockwise, alike $\varphi_0$.

The lines
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:qosset32}
  \State\textbf{\texttt{0.05;90;x+sqrt(4900+c1)-105}}\label{code:tee4}
  \State\textbf{\texttt{-sqrt(4900+c1)+105,147}}
  \State\textbf{\texttt{[-3*10\^{}3,0]}}
  \State\textbf{\texttt{[2,10]}}
  \State\textbf{\texttt{[32,448]}}\label{code:qosset42}
  \State\textbf{\texttt{0.0003;90;(x-sqrt(4900+c1)+105)\^{}2}}
  \textbf{\texttt{\hspace*{14ex}+(y-147)\^{}2-4900+c1}}\label{code:tee5}
  \State\textbf{\texttt{-3*sqrt(4900+c1)+105,147}}\label{code:triggerlast}
  \State\textbf{\texttt{[-3*10\^{}3,0]}}\label{code:teeset5}
\end{algorithmic}
defines the stages $\mathcal{M}_3$ and $\mathcal{M}_4$. Line~\ref{code:tee4} corresponds to TEE $\varphi_3(\mathbf{p}_k,c_{0,1}):=x+\sqrt{4900+c_{0,1}}-105$. Line~\ref{code:tee5} corresponds to TEE $\varphi_4(\mathbf{p}_k,c_{0,1}):=(x-\sqrt{4900+c_{0,1}}+105)^2+(y-147)^2-4900+c_{0,1}$.

Note that the survey mission example changes the radius of the circular TEE $\varphi_0$ through the parameter $c_{0,1}$. It also changes the radius of $\varphi_4$ in the same way. The other TEEs ($\varphi_1,\varphi_2,\varphi_3$) displacement changes the displacement altering the distance between the lines in the survey. A familiar pattern from Figure~\ref{fig:overview}. For simplicity, we suppose line~\ref{code:triggerlast} describe the last point of the mission. We further suppose the user does not desire to process any task in the last stage $\mathcal{M}_4$. To inhibit the computations
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:teeset5}
  \State\textbf{\texttt{[0,0]}}
  \State\textbf{\texttt{[0,0]}}
\end{algorithmic}

The following stages are constructed using the same principle. Th stages $\mathcal{M}_i$ for $i:=\{0,4,8,\dots\}$ contain TEE circle with variable radius. For $i:=\{1,3,5,\dots\}$ TEE line. For $i:=\{2,6,10,\dots\}$ TEE circle with fixed radius.

Once the reader is familiar with the initial mission plan format, a further abstraction is needed before we can simulate a mission in Appendix~\ref{app:simulator}. The algorithm needs to know how to model the tasks--the software components.

\section{Components Specification Format}

\section{Simulator}
\label{app:simulator}
